Robust Winner Determination and Simultaneous Elicitation of Scoring Rules and Preferences

### 2019-02-25 draft.tex
Strategies

\newif\ifappendix
%\appendixtrue %COMMENT to remove appendix % UNCOMMENT to add apendix
\input{ijcai_preamble} 
\author{ID: 2696}
%\author{Beatrice Napolitano$^1$\and Olivier Cailloux$^1$	\and  Paolo Viappiani$^{2}$
%	\affiliations $^1$ LAMSADE, UMR 7243, CNRS and Universit\'e Paris Dauphine, PSL Research University, Paris, France\\ $^2$ LIP6, UMR 7606, CNRS and Sorbonne Universit\'e, Paris, France
%	\emails \{firstname.lastname\}@dauphine.fr, paolo.viappiani@lip6.fr
%}
%Restoring page numbers as missing them out violates basic human rights of the reviewers. -- Olivier
\thispagestyle{plain}
\pagestyle{plain}

@@

\newcommand{\profile}{\bm{v}}%(complete) profile
\newcommand{\pprofile}{{\bm{p}}}%partial profile
\newcommand{\w}{\bm{w}}
\newcommand{\W}{\mathcal{W}}
\newcommand{\Co}{\mathcal{C}}
\newcommand{\pw}{W}%our knowledge about the weights
\newcommand{\powersetz}[1]{\mathscr{P}^*(#1)}

@@

\DeclarePairedDelimiter\set{\{}{\}}
\DeclarePairedDelimiter\card{\lvert}{\rvert}

@@

\begin{document}
\maketitle
\begin{abstract}
Social choice deals with the problem of determining a consensus choice from the preferences of different agents (voters).
In the classical setting, the voting rule is fixed beforehand and full preference information is provided.
Recently, the assumption of full preference information has been questioned by a number of researchers and several methods for eliciting preferences have been proposed.

In this paper we go one step further and we assume that both the voting rule and the agents preferences are partially specified.
Focusing on positional scoring rules, we assume that the chair, while  not able to give a precise definition of the rule, is capable of answering simple questions  requiring to pick a winner from a specific  example profile.
Moreover, the preferences of the agents are incrementally acquired by asking comparison queries.

We propose a method  for robust approximate winner determination in this setting with minimax regret. 
We then provide an interactive elicitation protocol based on minimax regret
and develop several query strategies that interleave questions to the chair and questions to the agents in order to acquire the most relevant information in order to quickly converge to optimal or a near-optimal alternative.
\end{abstract}

\section{Introduction}
%\commentOC{Do we agree that we should talk systematically about agents (or systematically about voters)? Currently we talk about voters in first section, then agents.}
Aggregation of preference information is a central task in many computer systems (recommender systems, search engines, etc).
In many situations, such as in group recommender systems, the goal is to find a consensus choice.
It is therefore natural to look at methods from social choice and see how they can be adapted for group decision-making in a computerized setting.

@@

%In this paper we propose an interactive adaptive protocol for eliciting both the agents preferences and the voting rule.
%At each step we need to decide whether we want to ask a question to the committee or to one of the agents (and to which agent in particular).
The paper is organized as follows.
In Section \ref{sec:background} we provide the necessary background.
We introduce the minimax criterion, that selects a winner that minimizes the worst possible regret, in Section \ref{sec:mmr}.
In Section \ref{sec:elicit} we provide an interactive elicitation protocol based on minimal regret;  in Section \ref{sec:experiments} we present the empirical validation of our approach with simulations; and Section \ref{sec:conclusions} provides some final thoughts.

\section{Social choice with partial information}
\label{sec:background}
We now introduce some basic concepts.
We assume defined a set $A$ of $m$ alternatives (products, restaurants, movies, public projects, job candidates, etc.) and a set $\set{1, …, n}$ of agents (voters); each agent $j$ is associated to her “real” preference order $\pref_j$. The agents come from an infinite set $\N$ of potential agents.
The real preferences of the agents are elements of $\linors$, the linear orders (connected, transitive, asymmetric relations) over the alternatives.
Following the social choice nomenclature, we call {\em profile} the association of a preference to each agent, considering a subset of agents from the set $\N$, and denote a profile by $(\pref_1,\ldots,\pref_n)$.
A profile is equivalently represented by $\profile=(v_1,\ldots,v_n)$ where $v_j(i) \in \set{1, \ldots, m}$ denotes the rank (position) of alternative $i$ in the preference order $\pref_j$.

Let $V$ be the set of possible preference profiles (the union for any integer $n$ of the $n$-fold cartesian product of the linear orders over the alternatives).
A social choice function $f : V \rightarrow \powersetz{A}$ associates a profile with a set of winners, where $\powersetz{A}$ represents the set of subsets of $A$ except for the emptyset. (Sets are used for tied winners.)
Among the many possible social choice functions, we consider {\em positional scoring rules (PSR)}, which attach weights to positions according to  the vector $(w_1, \ldots, w_m)$ (also called the scoring vector).
%A scoring rule associates each alternative to a score that is given by the sum of points obtained for each voter.
An alternative obtains a score that depends on the rank obtained in each of the preference orders:
\begin{align}
\label{eq:srule}
s(x; \profile, \w) = \sum_{j=1}^{n} w_{v_j(x)}
= \sum_{r=1}^{m} \alpha^{x}_r w_r 
\end{align}
where $\alpha^{x}_r$ is the number of times that alternative $x$ was ranked in the $r$-th position.
The winners are the alternatives with highest scores.
%We assume that the weights constitute a monotonic sequence: $w_{1} \geq w_{2} \geq \ldots \geq w_{m}$.

In this work we assume fixed, but unknown, a profile $\profile^*$ (and corresponding agents preferences $\pref_j^*$) and a weight vector $\w^*$, and we want to reason about partial preference information concerning those objects.
Our knowledge at a given time of the preference of voter $j$ is encoded by a partial order over the alternatives, thus a transitive and asymmetric binary relation, denoted by $\ppref_j$. 
In this work we assume that preference information is truthful, i.e. $a \ppref_j b ⇒ a \pref_j^* b$.
We use $\pinc$ to denote incomparability, that is $a \pinc_{j} b$ iff $a \nppref_j b \wedge b \nppref_j a$.
An incomplete profile $\pprofile = (\ppref_1, \ldots, \ppref_n)$ maps each voter to a partial preference.

@@

\medskip
We also assume that the weights of the scoring rule are only partially specified.
Therefore the vector $(w_1,\ldots,w_m)$  is not known but we are given a set of constraints restraining the possible values that weights can take.
We consider a decreasing sequence of weights\footnote{We assume all positions from $1$ to $m-1$ may contribute positively to the score of an alternative. The case where it is known that only the first $k<m-1$ positions matter can be easily dealt with by specifying $w_r = 0$ for $r > k$ as constraints. }:
% If the chair considers that, for example, “only the first ten positions matter” (so that $w_r = 0$ for $r > 10$), then we simply assume the chair is able to express this constraint at the start of the querying process, thus, that this information is known, and approximate the problem by considering that it is only about the first ten alternatives, thus, set $m=10$.}
\begin{align}
1=w_{1} > w_{2} > \ldots > w_{m}=0. \label{eq:monotone}
\end{align}
This is a natural assumption, as it is better to be ranked first than second, second than third, etc. 
Without loss of generality, we consider that $w_1=1$ and $w_m=0$. 
%\commentOC{I suggest we say no more than this to justify this assumption. Otherwise, we might say something like the following.

@@

Notice that if $\MMR(\pprofile, W)\!=\!0$, then $x^{*}_{\pprofile,W}$  is a necessary co-winner; this means that for any valid completion of the profile and any feasible $w \!\in\! W$, $x^{*}_{\pprofile,W}$ obtains a highest score.

@@

We are interested in computing $\PMR(x, y; \pprofile, W)$. This represents the “worst” difference of score, thus the difference of score between $y$ and $x$ under the worst case preferences compatible with $\pprofile$ and $W$, where the worst case is the one that maximizes this difference of score.
Consider our knowledge $\ppref_j$ of the preference of an agent $j$. 
The adversary has to make the score of $y$ as high as possible, and the score of $x$ as low as possible. 
To do this, he should complete $\ppref_j$ to $\pref_j$ by putting above $x$ as many alternatives as he can, that is, all the alternatives except those that are known to be worse than $x$ (those $a$ such that $x \ppref_j a$); and similarly, put below $y$ all the alternatives he can, thus, put above $y$ only those $a$ such that $a \ppref_j y$. 
Sometimes these two objectives conflict: when an alternative should go above $x$ according to the first objective and below $y$ according to the second one (because $¬(x \ppref_j a) ∧ ¬(a \ppref_j y)$), and $x$ is known to be better than $y$ (thus $x \ppref_j y$). 
In this case, the first objective should take priority, and $a$ should be put above $x$, which will move both $x$ and $y$ one rank lower than if $a$ had been put below $y$. 
This maximizes the adversary’s interests: because the weight vector is convex, the difference in scores will be lower when both alternatives are ranked lower (Equation \ref{eq:convexity}), and that difference of score is in favor of $x$ when $x \ppref_j y$, thus to be minimized from the point of view of the adversary.
To summarize:
\begin{align} 
\label{eq:compl}
\begin{split}
a \pref_j x &⇔ ¬(x \ppref_j a)\\
y \pref_j a &⇔ ¬(a \ppref_j y) ∧ (¬(x \ppref_j y) ∨ (x \ppref_j a)).
\end{split}
\end{align} 
In what follows let $\hat{\profile}$ be the profile computed according to the above procedure, and  $\hat{v}_j$ the $j$-th element (the linear order extending the partial order $p_j$ of the $j$-th voter).
We just sketched a proof of the following claim, and leave the rest of the proof to the reader.
\begin{claim}
There exists a completion $\hat{\profile} \in C(\pprofile)$ such that $\PMR(x,y; \pprofile, W) = \max_{\w \in W} [ s(y; \profile,\w) - s(x; \profile,\w) ]$ and that the linear order $\hat{v}_{j}$ of each agent $j$ satisfies (\ref{eq:compl}).
%There exists a completion ${\pref_j} \in C(\ppref_j)$ such that $\PMR(x,y; \pprofile, W) = \max_{\w \in W} \max_{\profile \in C(\pprofile), {\pref_j} \in \profile} [ s(y; \profile,\w) - s(x; \profile,\w) ]$ and that satisfies 
\end{claim}
%In this claim, we abuse notation by writing ${\pref_j} \in \profile$ to mean that the preference ordering corresponding to agent $j$ in $\profile$ is $\pref_j$.
\noindent It follows that $\hat{v}_{j}(x)$, the rank of $x$ for voter $j$ in the PMR-maximizing linear order $\bar{v}_{j}$ %$\PMR(x,y; \pprofile, W)$ 
is $\hat{v}_{j}(y) = 1+\card{A}-\card{{\ppref_j}(x)}$, where ${\ppref_j}(x)$ designates the set of alternatives that are less good than $x$ according to $\ppref_j$, and the rank of $y$ is $\hat{v}_{j}(x)=1+\card{{\pprefinv_j}(y)}+\beta$, where $\beta = \card{A} - \card{{\ppref_j}(x)}$ if $(x \ppref_j y)$ and $\beta = 0$ otherwise.

@@

\begin{claim}
The $\PMR$ can be written as:
\begin{align} 
\PMR(x,y; \pprofile, W)  
%\max_{w \in W} s(y; \hat{\profile}, \w) - s(y; \hat{\profile}, \w)\\
& = \max_{w \in W} \sum_{j=1}^n w_{\hat{v}_j(y)} - w_{\hat{v}_j(x)} = \\ 
& = \max_{w \in W} \sum_{r=1}^m (\hat{\alpha}_{r}^{y} - \hat{\alpha}_{r}^{x}) w_i. 
\end{align}
where the $\hat{\alpha}_{r}^{y}$ (resp. $\hat{\alpha}_{r}^{x}$)  is the number of times $y$ (resp. $x$) is at rank $r$ in the rankings of the completed profile $\hat{\profile}$.
%$\hat{\alpha}_{i}^{x} = \sum_{j=1}^{n}  I[\hat{v}_{j}(x)=i]$ where $I$ is the indicator function.
\end{claim}
%\[ \alpha_{i} = \sum_{j=1}^{n}  I[\hat{v}_{j}(y)=i] - I[\hat{v}_{j}(x)=i]\]
%We now show that the pairwise maximum regret can be computed with a linear program.
The PMR can be now computed by optimizing a linear program (LP) where the decision variables correspond to the weights attached to the different positions.
Constraints are on the feasible parameters (the weights of the scoring rules).
We remind that the preferences of the char are encoded with linear constraints $\Co_{W}$.
% contraints represent the preferences $\Co_{W}$ of the chair about the scoring rule.
The pairwise max regret $\PMR(x,y; \pprofile,W)$ is now obtained as the solution of the following linear program defined on the variables $w_1, …, w_m$. 
\begin{align*}
\max_{\w} & \sum_{r=1}^m (\hat{\alpha}_{r}^{y} - \hat{\alpha}_{r}^{x}) w_{r}\\
\text{ s.t. } & \text{Equation } (\ref{eq:monotone})\\
& \text{Equation } (\ref{eq:convexity})\\
& \Co_W.
\end{align*}
Note that given our choice $w_{1}=1$ and $w_{m}=0$, there are only $m-2$ variables 
(we leave $w_{1}$ and $w_{m}$ in the LP just for clarity of presentation).

%Finally the minimax regret solution $\MMR(\pprofile, W)$ is determined by solving a sequence of pairwise max regret computations, solving the above linear program for every different pair $(x,y) \in A^{2}$.
The max regret $\MR(x; \pprofile, W)$ is determined by considering the pairwise regret of $x$ with each other alternative in $A$.
The optimal alternative w.r.t. minimax regret is the one with least max regret. 
Notice that, whenever the PMR of an alternative $x$ (against some other alternative) exceeds the best MR value found so far, we don't need to further evaluate $x$. 
This idea can be exploited further by adopting a minimax-search tree (see \cite{Braziunas2011}).
%The computation can be made faster by adopting an efficient scheme of evaluation of PMR where .

@@

The remainder of this Section is structured as follows.
First of all, we discuss the different types of questions that can be asked to the agents and to the chair, and the way responses are handled.
Then, we describe different strategies to determine informative queries to ask next, with the goal of reducing $\MMR(\pprofile,W)$ quickly.

%Type of questions that we can ask to the chair:	bound queries (Is this alternative among your top-k most preferred items?), comparison queries (do you prefer alternative x or alternative y?).
% we discuss the different forms of queries and develop several query strategies 
\paragraph{Question types}
We distinguish between questions asked to the agents and questions asked to the chair.
As {\em questions asked to the agents} it is natural to consider comparison queries asking to compare two alternatives.
%Another common type of queries are {\em top-k}, asking to each voter her $k$ most preferred alternatives.
The effect of a response to a question asked to an agent is to increase our knowledge about the agents rankings, thus augmenting the partial profile $\pprofile$. 
If voter $j$ answers a comparison query stating that alternative $a$ is preferred to $b$, then the partial order $\ppref_j$ is augmented with $a \ppref_j b$ and by transitive closure.

A bit more discussion is needed about {\em questions asked to the chair}.
Such questions aim at refining our knowledge about the scoring rule; a response gives us a constraint on the weight vector $\w$.
In particular, we want to acquire constraints of the type:
\[ w_{r} - w_{r+1} \geq \lambda (w_{r+1} - w_{r+2}) \] 
for $r \in \{1,\ldots,m-2\}$, relating the difference between the importance of ranks $r$ and $r+1$ with the difference between ranks $r+1$ and $r+2$.

\paragraph{Building concrete questions for the chair}
As we assume the weights constitute the utility components of the chair, it might be reasonable to assume, 
%as we do here, 
that the chair is able to answer such abstract questions in our setting. However, it is important to make sure that the question can also, in principle, be asked in a more concrete way, in terms of winners of example profiles. This permits to test whether the chair understands the question as we do; this permits to relate the preference of the chair  to her choice behavior in the economic sense; and this will be necessary for an ordinal extension of our work where the scores would not be considered as cardinal utilities.
Thus, our task is to build a profile, given $\lambda$ and $r ≤ m-2$, in such a way that the set of (tied) winners picked by the chair reveals whether $w_{r} - w_{r+1} \geq \lambda (w_{r+1} - w_{r+2})$.
We assume that $\lambda$ is rational, and define $\lambda = p/q$.

Observe that the question may be defined equivalently as whether $q w_{r} + p w_{r+2} ≥ (p + q) w_{r+1}$, where $p, q$ are natural numbers. 
As a first attempt to make this question concrete, let us define a profile $P$ containing $p+q$ agents. 
Define an alternative $a$ that receives $q$ times the rank $r$ and $p$ times the rank $r+2$ and an alternative $b$ that receives $p+q$ times the rank $r+1$. 
Observing that the score of $a$ in that profile is exactly the left hand side of the question, and that the score of $b$ is the right hand side, intuition suggests that observing whether the chair picks $a$ or $b$ as winner will let us determine which side is greater; equality occurring when the chair declares $a$ and $b$ as tied for the victory. 
However, we still need to complete the profile, thus, come up with other $m-2$ alternatives defined so that each agent in the resulting profile has placed exactly one alternative in each rank. 
And we must ensure, doing this, that these other alternatives are not better than $a$ or $b$.\footnote{Because, if the chair picks a different alternative $c$, it tells us that the chair prefers $c$ to both $a$ and $b$, 
%but it generally does not  answer the question we are interested in
but in general we do not acquire the type of constraint we are interested in.} 
Taking $p = q = 1$, $m=4$, $r=2$, we see that such a completion may be impossible. Luckily, this problem can be worked around, at the price of increasing the number of agents.

First build a temporary profile $P$ of $p+q$ agents with $a$ and $b$ ranked as just described. Complete it with $m-2$ alternatives ranked in arbitrary orders so that the resulting profile has complete strict rankings for each of the $p+q$ voter. Now we will make the other alternatives as bad as desired by adding agents to $P$ that appreciate $a$ and $b$ more than the other alternatives, thus building a profile $P'$ in which, whatever the weights, $c$ may not have a better score than $a$. Observe that if we add $\delta$ agents (with $\delta$ a natural number) that put $a$ at first rank and $b$ at second rank, and $\delta$ agents that put $b$ at first rank and $a$ at second rank, we do not change the difference of the scores of $a$ and $b$, and thus, observing the chair choosing $a$ or $b$ still answers the question. Choose arbitrarily a matching of the remaining $m-2$ alternatives to the ranks $3$ to $m$, say, $c \mapsto 3$, $d \mapsto 4$, and so on. The first $\delta$ supplementary agents, who rank $a$ first and $b$ second, rank the other alternatives as the just defined mapping suggests. The other $\delta$ supplementary agents, who rank $b$ first and $a$ second, use the opposite ranking for the remaining alternatives, say, $c \mapsto m$, $d \mapsto m-1$, and so on. We can now prove that, for $\delta = p+q$, and whatever the weight vector $w \in \W$, no alternative, except possibly $b$, has a better score than $a$. (In fact, as the construction will exhibit, they are all worst than both $a$ and $b$, but the weaker fact is enough for our claim.)

>>>>

\newif\ifappendix
\appendixtrue %COMMENT to remove appendix % UNCOMMENT to add apendix
\input{ijcai_preamble} 
\author{Beatrice Napolitano$^1$\and Olivier Cailloux$^1$	\and  Paolo Viappiani$^{2}$
	\affiliations $^1$ LAMSADE, UMR 7243, CNRS and Universit\'e Paris Dauphine, PSL Research University, Paris, France\\ $^2$ LIP6, UMR 7606, CNRS and Sorbonne Universit\'e, Paris, France
	\emails \{firstname.lastname\}@dauphine.fr, paolo.viappiani@lip6.fr
}
%Restoring page numbers as missing them out violates basic human rights of the reviewers. -- Olivier
\thispagestyle{plain}
\pagestyle{plain}

@@

\newcommand{\profile}{\bm{v}}%(complete) profile
\newcommand{\pprofile}{{\bm{p}}}%partial profile
\newcommand{\w}{\bm{w}}
\newcommand{\W}{\mathcal{W}}
\newcommand{\pw}{W}%our knowledge about the weights
\newcommand{\powersetz}[1]{\mathscr{P}^*(#1)}
\newcommand{\strat}[1]{\emph{#1}}

@@

\DeclarePairedDelimiter\set{\{}{\}}
\DeclarePairedDelimiter\card{\lvert}{\rvert}
\DeclarePairedDelimiter\abs{\lvert}{\rvert}

@@

\begin{document}
\maketitle
\begin{abstract}
Social choice deals with the problem of determining a consensus choice from the preferences of different agents (voters).
In the classical setting, the voting rule is fixed beforehand and full preference information is provided by the voters. 
Recently, the assumption of full preference information has been questioned by a number of researchers and several methods for eliciting voters information have been proposed.

In this paper we go one step further and we assume that both the voting rule and the voters preferences are partially specified.
Focusing on positional scoring rules, we assume that the chair, while  not able to give a precise definition of the rule, is capable of answering simple questions  requiring to pick a winner from a specific  example profile.
Moreover, the preferences of the voters are incrementally acquired by asking comparison queries.

We propose a method  for robust approximate winner determination in this setting with minimax regret. 
We then provide an interactive elicitation protocol based on minimax regret
and develop several query strategies that interleave questions to the chair and questions to the voters in order to acquire the most relevant information in order to quickly converge to optimal or a near-optimal alternative.
\end{abstract}

\section{Introduction}
\commentOC{Do we agree that we should talk systematically about agents (or systematically about voters)? Currently we talk about voters in first section, then agents.}
Aggregation of preference information is a central task in many computer systems (recommender systems, search engines, etc).
In many situations, such as in group recommender systems, the goal is to find a consensus choice.
It is therefore natural to look at methods from social choice and see how they can be adapted for group decision-making in a computerized setting.

@@

%In this paper we propose an interactive adaptive protocol for eliciting both the voters preferences and the voting rule.
%At each step we need to decide whether we want to ask a question to the committee or to one of the agents (and to which agent in particular).
The paper is organized as follows.
In Section \ref{sec:background} we provide the necessary background.
We introduce the minimax criterion, that selects a winner that minimizes the worst possible regret, in Section \ref{sec:mmr}.
In Section \ref{sec:elicit} we provide an interactive elicitation protocol based on minimal regret;  in Section \ref{sec:experiments} we present the empirical validation of our approach with simulations; and Section \ref{sec:conclusions} provides some final thoughts.

\section{Social choice with partial information}
\label{sec:background}
We now introduce some basic concepts.
We assume defined a set $A$ of $m$ alternatives (products, restaurants, movies, public projects, job candidates, etc.) and a set $\set{1, …, n}$ of agents (voters); each agent $j$ is associated to her “real” preference order $\pref_j$. The agents come from an infinite set $\N$ of potential voters.
The real preferences of the agents are elements of $\linors$, the linear orders (connected, transitive, asymmetric relations) over the alternatives.
Following the social choice nomenclature, we call {\em profile} the association of a preference to each voter, considering a subset of voters from the set $\N$, and denote a profile by $(\pref_1,\ldots,\pref_n)$.
A profile is equivalently represented by $\profile=(v_1,\ldots,v_n)$ where $v_j(i) \in \set{1, \ldots, m}$ denotes the rank (position) of alternative $i$ in the preference order $\pref_j$.

Let $V$ be the set of possible preference profiles (the union for any integer $n$ of the $n$-fold cartesian product of the linear orders over the alternatives).
A social choice function $f : V \rightarrow \powersetz{A}$ associates a profile with a set of winners, where $\powersetz{A}$ represents the set of subsets of $A$ except for the emptyset. (Sets are used for tied winners.)
Among the many possible social choice functions, we consider {\em positional scoring rules}, which attach weights to positions according to  the vector $(w_1, \ldots, w_m)$ (also called the scoring vector).
%A scoring rule associates each alternative to a score that is given by the sum of points obtained for each voter.
An alternative obtains a score that depends on the rank obtained in each of the preference orders:
\begin{align}
\label{eq:srule}
s(x; \profile, \w) = \sum_{j=1}^{n} w_{v_j(x)}
= \sum_{r=1}^{m} \alpha^{x}_r w_r 
\end{align}
where $\alpha^{x}_r$ is the number of times that alternative $x$ was ranked in the $r$-th position.
The winners are the alternatives with highest scores.
%We assume that the weights constitute a monotonic sequence: $w_{1} \geq w_{2} \geq \ldots \geq w_{m}$.

In this work we assume fixed, but unknown, a profile $\profile^*$ (and corresponding voters preferences $\pref_j^*$) and a weight vector $\w^*$, and we want to reason about partial preference information concerning those objects.
Our knowledge at a given time of the preference of voter $j$ is encoded by a partial order over the alternatives, thus a transitive and asymmetric binary relation, denoted by $\ppref_j$. 
In this work we assume that preference information is truthful, i.e. $a \ppref_j b ⇒ a \pref_j^* b$.
We use $\pinc$ to denote incomparability, that is $a \pinc_{j} b$ iff $a \nppref_j b \wedge b \nppref_j a$.
An incomplete profile $\pprofile = (\ppref_1, \ldots, \ppref_n)$ maps each voter to a partial preference.

@@

\medskip
We also assume that the weights of the scoring rule are only partially specified.
Therefore the vector $(w_1,\ldots,w_m)$  is not known but we are given a set of constraints restraining the possible values that weights can take.
We consider a decreasing sequence of weights:
\begin{align}
1=w_{1} > w_{2} > \ldots > w_{m}=0. \label{eq:monotone}
\end{align}
This is a natural assumption, as it is better to be ranked first than second, second than third, etc. 
Without loss of generality, we consider that $w_1=1$ and $w_m=0$. \commentOC{I suggest we say no more than this to justify this assumption. Otherwise, we might say something like the following. If the chair considers that, for example, “only the first ten positions matter” (so that $w_r = 0$ for $r > 10$), then we simply assume the chair is able to express this constraint at the start of the querying process, thus, that this information is known, and approximate the problem by considering that it is only about the first ten alternatives, thus, set $m=10$.}

@@

Notice that if $MMR(\pprofile, W)=0$, then $x^{*}_{\pprofile,W}$  is a necessary co-winner; this means that for any valid completition of the profile and any feasible weight $w \in W$, $x^{*}_{\pprofile,W}$ obtains a highest score.

@@

We are interested in computing $\PMR(x, y; \pprofile, W)$. This represents the “worst” difference of score, thus the difference of score between $y$ and $x$ under the worst case preferences compatible with $\pprofile$ and $W$, where the worst case is the one that maximizes this difference of score.
Consider our knowledge $\ppref_j$ of the preference of an agent $j$. The adversary has to make the score of $y$ as high as possible, and the score of $x$ as low as possible. To do this, he should complete $\ppref_j$ to $\pref_j$ by putting above $x$ as many alternatives as he can, that is, all the alternatives except those that are known to be worst than $x$ (those $a$ such that $x \ppref_j a$); and similarly, put below $y$ all the alternatives he can, thus, put above $y$ only those $a$ such that $a \ppref_j y$. Sometimes these two objectives conflict: when an alternative should go above $x$ according to the first objective and below $y$ according to the second one (because $¬(x \ppref_j a) ∧ ¬(a \ppref_j y)$), and $x$ is known to be better than $y$ (thus $x \ppref_j y$). In this case, the first objective should take priority, and $a$ should be put above $x$, which will move both $x$ and $y$ one rank lower than if $a$ had been put below $y$. This maximizes the adversary’s interests: because the weight vector is convex, the difference in scores will be lower when both alternatives are one rank lower (Equation \ref{eq:convexity}), and that difference of score is in favor of $x$ when $x \ppref_j y$, thus to be minimized from the point of view of the adversary.
To summarize:\commentOC{TODO only one number here} 
\begin{align} 
\label{eq:compl}
a \pref_j x &⇔ ¬(x \ppref_j a)\\
y \pref_j a &⇔ ¬(a \ppref_j y) ∧ (¬(x \ppref_j y) ∨ (x \ppref_j a)).
\end{align}

We just sketched a proof of the following claim, and leave the rest of the proof to the reader.
\begin{claim}
	There exists a completion ${\pref_j} \in C(\ppref_j)$ such that $\PMR(x,y; \pprofile, W) = \max_{\w \in W} \max_{\profile \in C(\pprofile), {\pref_j} \in \profile} [ s(y; \profile,\w) - s(x; \profile,\w) ]$ and that satisfies \ref{eq:compl}.
\end{claim}
In this claim, we abuse notation by writing ${\pref_j} \in \profile$ to mean that the preference ordering corresponding to agent $j$ in $\profile$ is $\pref_j$.

It follows that the rank of $x$ for voter $j$ in $\PMR(x,y; \pprofile, W)$ is $1+\card{A}-\card{{\ppref_j}(x)}$, where ${\ppref_j}(x)$ designates the set of alternatives that are less good than $x$ according to $\ppref_j$, and the rank of $y$ is $1+\card{{\pprefinv_j}(y)}+\alpha$, where $\alpha = \card{A} - \card{{\ppref_j}(x)}$ if $(x \ppref_j y)$ and $\alpha = 0$ otherwise.

@@

The space of feasible weights $W$ is encoded with linear constraints that model the preferences of the chair.
Therefore the pairwise maximum regret can be computed with a linear program.

The minimax regret solution $\MMR(\pprofile, W)$ is determined by solving a sequence of pairwise max regret computations.
Each PMR is computed using a linear program.
The decision variables correspond to the weights attached to the different positions.
Constraints are on the feasible parameters (the weights of the scoring rules); contraints represent the preferences of the chair about the scoring rule.

Let $\hat{\profile}$ be the profile computed according to the above procedure, and  $\hat{v}_j$ the $j$-th element (the linear order extending the partial order $p_j$ of the $j$-th voter).
\begin{prop}
%Assume convex weights.
The $\PMR$ can be written as:
\begin{align} 
\PMR(x,y; \pprofile, W) =  
\max_{w \in W} s(y; \hat{\profile}, \w) - s(y; \hat{\profile}, \w)\\
=\max_{w \in W} \sum_{j=1}^n w_{\hat{v}_j(y)} - w_{\hat{v}_j(x)} =
\max_{w \in W} \sum_{r=1}^m (\hat{\alpha}_{r}^{y} - \hat{\alpha}_{r}^{x}) w_i 
\end{align}
where the coefficient $ (\hat{\alpha}_{r}^{y} - \hat{\alpha}_{r}^{x})$ is the number of times $y$ is at rank $r$ minus the number of times $x$ is at rank $r$ in the rankings of the completed profile $\hat{\profile}$.
%$\hat{\alpha}_{i}^{x} = \sum_{j=1}^{n}  I[\hat{v}_{j}(x)=i]$ where $I$ is the indicator function.
\end{prop}
%\[ \alpha_{i} = \sum_{j=1}^{n}  I[\hat{v}_{j}(y)=i] - I[\hat{v}_{j}(x)=i]\]
The pairwise max regret $\PMR(x,y; \pprofile,W)$ is now obtained as the solution of the following linear program defined on the variables $w_1, …, w_m$. 
\begin{align*}
\max_{\w} & \sum_{r=1}^m (\hat{\alpha}_{r}^{y} - \hat{\alpha}_{r}^{x}) w_{r}\\
\text{ s.t. } & \text{Equation } (\ref{eq:monotone})\\
& \text{Equation } (\ref{eq:convexity})\\
& C_W.
\end{align*}
Note that given our choice $w_{1}=1$ and $w_{m}=0$, there are only $m-2$ variables. We leave $w_{1}$ and $w_{m}$ in the linear program just for clarity of presentation.

@@

The remainder of this Section is structured as follows.
First of all, we discuss the different types of questions that can be asked to the voters and to the chair, and the way responses are handled.
Then, we describe different strategies to determine informative queries to ask next, with the goal of reducing $\MMR(\pprofile,W)$ quickly.

\paragraph{Question types}
We distinguish between questions asked to the voters and questions asked to the chair.
As {\em questions asked to the voters}, it is natural to consider comparison queries asking to compare two alternatives.
%Another common type of queries are {\em top-k}, asking to each voter her $k$ most preferred alternatives.
The effect of a response to a question asked to a voter is to augment our knowledge about the voters rankings, thus augmenting the partial profile $\pprofile$. 
If voter $j$ answers a comparison query stating that alternative $a$ is preferred to $b$, then the partial order $\ppref_j$ is augmented with $a \ppref_j b$ and by transitive closure.

A bit more discussion is needed about {\em questions asked to the chair}.
Such a question aims at refining our knowledge about the scoring rule;
a response gives us a constraint on the weight vector $\w$.
In particular, we want to acquire constraints of the type:
\[ w_{r} - w_{r+1} \geq \lambda (w_{r+1} - w_{r+2}) \] 
for $r \in \{1,\ldots,m-2\}$, relating the difference between the importance of ranks $r$ and $r+1$ with the difference between ranks $r+1$ and $r+2$.

\paragraph{Building concrete questions for the chair}
As we assume the weights constitute the utility components of the chair, it might be reasonable to assume, as we do here, that the chair is able to answer such abstract questions in our setting. However, it is important to make sure that the question can also, in principle, be asked in a more concrete way, in terms of winners of example profiles. This permits to test whether the chair understands the question as we do; this permits to relate the preference of the chair as we understand it to her choice behavior in the economic sense; and this will be necessary for an ordinal extension of our work where the scores would not be considered as cardinal utilities.
Thus, our task is to build a profile, given $\lambda$ and $r ≤ m-2$, in such a way that the set of (tied) winners picked by the chair reveals whether $w_{r} - w_{r+1} \geq \lambda (w_{r+1} - w_{r+2})$.
We assume that $\lambda$ is rational, and define $\lambda = p/q$.

Observe that the question may be defined equivalently as whether $q w_{r} + p w_{r+2} ≥ (p + q) w_{r+1}$, where $p, q$ are natural numbers. As a first attempt to make this question concrete, let us define a profile $P$ containing $p+q$ voters. Define an alternative $a$ that receives $q$ times the rank $r$ and $p$ times the rank $r+2$ and an alternative $b$ that receives $p+q$ times the rank $r+1$. Observing that the score of $a$ in that profile is exactly the left hand side of the question, and that the score of $b$ is the right hand side, intuition suggests that observing whether the chair picks $a$ or $b$ as winner will let us determine which side is greater; equality occurring when the chair declares $a$ and $b$ as tied for the victory. However, we still need to complete the profile, thus, come up with other $m-2$ alternatives defined so that each voter in the resulting profile has given each rank once. And we must ensure, doing this, that these other alternatives are not better than $a$ or $b$. For if the chair picks a different alternative $c$, it only tells us that the chair prefers $c$ to both $a$ and $b$, but it generally does not answer the question we are interested in. Taking $p = q = 1$, $m=4$, $r=2$, we see that such a completion may be impossible. Luckily, this problem can be worked around, at the price of increasing the number of voters.

First build a temporary profile $P$ of $p+q$ voters with $a$ and $b$ ranked as just described. Complete it with $m-2$ alternatives ranked in arbitrary orders so that the resulting profile has complete strict rankings for each of the $p+q$ voter. Now we will make the other alternatives as bad as desired by adding voters to $P$ that appreciate $a$ and $b$ more than the other alternatives, thus building a profile $P'$ in which, whatever the weights, $c$ may not have a better score than $a$. Observe that if we add $\delta$ voters (with $\delta$ a natural number) that put $a$ at first rank and $b$ at second rank, and $\delta$ voters that put $b$ at first rank and $a$ at second rank, we do not change the difference of the scores of $a$ and $b$, and thus, observing the chair choosing $a$ or $b$ still answers the question. Choose arbitrarily a matching of the remaining $m-2$ alternatives to the ranks $3$ to $m$, say, $c \mapsto 3$, $d \mapsto 4$, and so on. The first $\delta$ supplementary voters, who rank $a$ first and $b$ second, rank the other alternatives as the just defined mapping suggests. The other $\delta$ supplementary voters, who rank $b$ first and $a$ second, use the opposite ranking for the remaining alternatives, say, $c \mapsto m$, $d \mapsto m-1$, and so on. We can now prove that, for $\delta = p+q$, and whatever the weight vector $\w \in \W$, no alternative, except possibly $b$, has a better score than $a$. (In fact, as the construction will exhibit, they are all worst than both $a$ and $b$, but the weaker fact is enough for our claim.)
